---
title: "Rooting My Way Through Functions"
author: "Ethan"
date: "2020-01-27"
output: html_document
---

# The Setup

Welcome to another wonderful installment in "(Ou)R Adventures with Ethan". Today we will be doing something quite interesting and fun: writing a function that can find the square root of a number you give it.

Now, you may be thinking, *"What's the point of this? R already has a built in function to find the square root of a number."* But, you see, writing your own function from scratch that can find the square root of a number, will not only provide you with practice in constructing functions, but also give you a chance to exercise those (hopefully not atrophied) algebra skills.  

# The Function

Here we have the function in all of its glory. I have included the explanation for each line in the comments within the code:

```{r}
mysqrt <- function(x, myguess=10, tol=.000001, verbose=FALSE) { #This sets the default parameters of the function
	x <- if(x < 0){return("Stop! That's illegal. This function will not allow you to take the square root of a negative number")} else if(x==0) {return(0)} else{x} #This line filters the x input to only allow the function to take square roots of postive numbers. If the input passes this test, then it the function will determine if the input is "0". If the input is "0" then the function will return "0" becuase the square root of "0" is "0". If the input is greater than "0", the input will be stored as "x". 
	newguess <- if(myguess < 0){return("Stop! That's illegal. The square root of a number cannot be negative")} else{myguess} #This line filters the inputs to only allow the function to accept guesses that are postive numbers. 
	approxy <- x/newguess 
	newguess <- (newguess + approxy)/2
	error <- newguess-approxy
	error <- if(error < 0) {error*-1} else if (error == 0){return("Your guess is correct!")} else{error} #This line is a fun conditional. If your guess is below the square root of "x" you will get a negative number for "error". In order to keep the error postive for the logical argument "error > tol" to work, the function will mulitply all negative "error" values by "-1". If your guess is correct, it will return a message stating that your guess was correct!
	error < tol
	if(verbose==TRUE) {print(paste("error:", error))} #This is a conditional that prints the difference between newguess and approxy for the first round if verbose==TRUE
	if(verbose==TRUE) {print(paste("approximate y:", approxy))} #This is a conditional the prints the approxy from the algorithm above for the first round if verbose==TRUE
	if(verbose==TRUE) {print(paste("guess Y:", newguess))} #This is a conditional the prints the new guess from the algorithm above for the first round if verbose==TRUE

	while(error > tol) { #This is the while loop that will recalculate the "error" value until it is under the tolerance specifed in the argument "tol".
		approxy <- x/newguess
		newguess <- (newguess + approxy)/2 
		error <- newguess-approxy
		if(verbose==TRUE) {print(paste("error:", error))}  #This is a conditional that prints the difference between newguess and approxy for each iteration of the "while loop" if verbose==TRUE
		if(verbose==TRUE) {print(paste("approximate y:", approxy))} #This is a conditional the prints the approxy from the algorithm above for each iteration of the "while loop" if verbose==TRUE
		if(verbose==TRUE) {print(paste("guess Y:", newguess))} #This is a conditional the prints the new guess from the algorithm above for for each iteration of the "while loop" if verbose==TRUE
}
	if(verbose==TRUE) {return(paste("final answer:", round(newguess, 3)))} else{return(newguess)} #This is the "return" line of the function. If verbose==T, it will mark the "newguess" that is associated with an error < tolerance as "final answer" and round it to 3 decimal places. If verbose==F, it will just return the "newguess" that is associated with an error < tolerance.
}
```
# The Function in Action

The "x" argument is number you want to square root, and the default behavior of the argument "verbose" is set to FALSE:
```{r}
mysqrt(x=4)
```

This will also work because I have set x as the first argument:
```{r}
mysqrt(4)
```

The x variable is number you want to square root and you can set the argument verbose=T to show the steps it takes to reach that answer:
```{r}
mysqrt(x=4, verbose=T)
```

Now you can try it on some random number:
```{r}
mysqrt(1234567)
```

The function will work if your guess is greater than the "x" argument:
```{r}
mysqrt(x=64, myguess=100)
```

The function will work if your guess is less than your "x" argument:
```{r}
mysqrt(x=64, myguess=1)
```

The function will print a fun message if you guess the square root correctly:
```{r}
mysqrt(x=64, myguess=8)
```

The function will print a fun message if your guess is a negative number:
```{r}
mysqrt(x=64, myguess=-15)
```

The function will print a fun message if you try to take the square root of a negative number:
```{r}
mysqrt(x=-16)
```

# Further Testing
You can actually try this yourself! If you run the function in your R console, you can continue playing with it! 

Thank you for tuning in for this installment of "(Ou)R Adventures with Ethan"!